---
title: "ARA 669 - Assignment 1"
author: "Your name here"
date: "today's date here"
output: html_document
---

## Assignment

This first assignment is designed to ease you into the course, and to introduce GitHub and Quarto to those of you not familiar with the programs. It's also an opportunity for me to gauge a little more formally where everyone is in terms of coding styles and experiences. 

Please submit this assignment via GitHub by Sept. 3 at 5 pm.

## Getting started

Use the code chunk below to load the tidyverse package \[0.5 pt\].

```{r setup}
#| message: false
#| error: false
installed.packages("tidyverse)
library(tidyverse)
```
## Data exploration
Several well-behaved datasets are preloaded into R and are frequently used to demonstrate statistical concepts. Use the function data() to retrieve a list available datasets and pick one to explore for this assignment.

```{r explore}
#| message: false
#| error: false
data("women")
```
Describe the context of the dataset and the structure of the data themselves \[1 pt\].
This data includes the height adn weight of 15 women 
## Plotting
Explore the data visually using ggplot2 \[1 pt\].

```{r plotting}
#| message: false
#| error: false

installs.packages(dplyr)
library(dplyr)

df <- airquality %>%
  mutate(
    date = as.Date(paste(1973, Month, Day, sep = "-"), format = "%Y-%m-%d")
  )
  
ggplot(data = df, mapping = aes(date, Temp ))+geom_point()
ggplot(data = df, mapping = aes(date, Temp, colour = Solar.R ))+geom_point()

```
Describe your approach. What choices did you make in deciding which variables to use, plot types, etc., and why? [1 pt].
This data set included two columns of month and day instead of a date column. When I went to plot temperature as a function of time, I got very weird outcomes because month was 5-9 and day was 1-30. So first I created a "date" column so that it would appear as year, month, day. Then I plotted time as x and temperature as y in ggplot. I added Solar radiation as the colour scheme. 

Read the perspectives article “Should we still teach or learn coding? A postgraduate student perspective on the use of large language models for coding in ecology & evolution” (Campbell et al. 2024). What part(s) of this article resonated with you, and why [1.5 pts]? 

There is a lot that resonated with me in the limitations of LLMs section in this article. I think it is very important to be cautious with the outputs you are given because they are not always right. At the same time, I think LLMs are best used for coding...to an extent! They are a great source to troubleshoot your code. And you know it provides a "right" answer if your code runs. This is assignment is a great example. I knew I wanted to plot temperature as a function of time and I understood my dataset in that it lacked a proper "date" column. I used descriptions of R packages as much as I could to construct this new date column. Once I got stuck, I used claude code to help explain where and why the function would not work. I used claude code as a tool behind my code. I also agree that coding must still be taught in school. Coding is like learning a language, it is vital to understand the backbone and basics of the language before you fold in complicated commands, loops, etc. Lastlty, I never considered LLMs to be "more nice" than coding websites and online platforms. While some may consider this a beneift of LLMs, I see it as a tool so the "niceness" of LLMs is redundant...and almost creepy. 

## AI statement

Describe your use of AI in this assignment, including:

1.  What tool or service did you use? I used claude code to help me with the mutate function in dplyr
2.  What was the specific purpose or use? To understand how to execute commands like ggplot in R studio, I look at the packages information in the bottom right corner. Sometimes even though I can read the guidelines of the commands, I still cannot figure out how to fill them. When I need to create a new column called "date" I used dplyr and mutate. I kept getting an error until Claude told me to include "format = %Y-%m-%d" and it finally worked.
3.  How did this use enhance your understanding of the topic/code? I could not get a new column to format propoerly so now I know how to do that in the future. And specifically how to create date columns which is commonly used in stats! 

